Sparse Function-space Representation \\ of Neural Networks

### 2023-05-24 paper/main.tex
commented the table
update app A

In binary classification, we deal with Bernoulli likelihoods. The inverse link function is the sigmoid function $h(f) = \sigma(f) = \frac{1}{1 + e^{-f}}$, and we can then write the negative log likelihood as
\begin{equation}
\begin{aligned}
	\ell(h(f), y) 
	&= -\log \left( \sigma(f)^y (1 - \sigma(f))^{1-y} \right) \\
	&= - y \log(\sigma(f)) - (1-y) \log(1- \sigma(f)) \\
	&= - y f + \log(1 + e^f) 
\end{aligned}
\end{equation}
Then, it follows that the the derivative we are interested in are
\begin{equation}
	\nabla_f \ell(\sigma(f), y_i)|_{f=f_i} = - y_i + \sigma(f_i) \quad \text{ and } \quad  \nabla_{f\!f} \ell(h(f), y_i)|_{f=f_i} = \sigma(f_i) (1 - \sigma(f_i)).
\end{equation}

Moving to the multi class classification, the function outputs is no more a scalar but a vector of logits $\bm{f} \in \R^C$. The inverse link function for the categorical loss is the softmax function $\textrm{Softmax}(\bm{f}) : \R^C \to \R^C$ defined as $h(\bm{f}) = \textrm{Softmax}(\bm{f}) = \frac{\exp(\bm{f})}{ \sum^C_{j=1} \exp(\bm{f}_j)} \in \R^C$. If we write the targets $y_i$ as a one-hot $C$-shaped one-hot encoding vector, we can write the negative log likelihood as:
\begin{equation}
\begin{aligned}
	\ell(h(\bm{f}),\vy) &= - \log\left( \vy^\T \textrm{Softmax}(\bm{f})\right)\\
	&= - y_i \log \left( \left[ \textrm{Softmax}(\bm{f}) \right]_i \right)
\end{aligned}
\end{equation}
where $\left[\textrm{Softmax}(\bm{f})\right]_i = \frac{\exp(\bm{f}_i)}{\sum_{j=1}^{C} \exp(\bm{f}_j)}$ defines the $i$\textsuperscript{th} entry of the logits fed through the $\textrm{Softmax}(\cdot)$ function.

@@

\begin{table}[!ht]
\centering
\caption{}
\begin{tabular}{@{}cccc@{}}
\toprule
Loss        & $h(\cdot)$ & $\nabla_{f}$ & $\nabla^2_{f\!f}$  \\ \midrule
Bernoulli   & $\sigma(f) = \frac{1}{1 + e^{-f}}$ & $\sigma(f) - y$   & $\sigma(f)(1-\sigma(f))$ \\
MSE         & & $y - f$      & $1$ \\
Categorical & $\mathcal{S}_i(\vf) = \frac{e^{\vf_i}}{\sum_{j=1}^C e^{\vf_j}}$ &  $\begin{bmatrix}
		p_1\\ \vdots\\p_k
	\end{bmatrix}$     &   $\begin{bmatrix}
		p_1\\ \vdots\\p_k
	\end{bmatrix}$ \\ \bottomrule
\end{tabular}
\end{table}

>>>>

In binary classification, we deal with Bernoulli likelihoods. The inverse link function is the sigmoid function $h(f) = \sigma(f) = \frac{1}{1 + e^{-f}}$, and we can then write the negative log likelihood as
\begin{equation}
\begin{aligned}
	\ell(h(f), y) 
	&= -\log \left( \sigma(f)^y (1 - \sigma(f))^{1-y} \right) \\
	&= - y \log(\sigma(f)) - (1-y) \log(1- \sigma(f)) \\
	&= - y f + \log(1 + e^f) 
\end{aligned}
\end{equation}
Then, it follows that the the derivative we are interested in are
\begin{equation}
	\nabla_f \ell(h(f), y_i)|_{f=f_i} = - y_i + \sigma(f_i) \quad \text{ and } \quad  \nabla_{f\!f} \ell(h(f), y_i)|_{f=f_i} = \sigma(f_i) (1 - \sigma(f_i)).
\end{equation}

Moving to the multi class classification, the function outputs is no more a scalar but a vector of logits $\bm{f} \in \R^C$. The inverse link function for the categorical loss is the softmax function $\textrm{Softmax}(\bm{f}) : \R^C \mapsto \R^C$ defined as $h(\bm{f}) = \textrm{Softmax}(\bm{f}) = \frac{\exp(\bm{f})}{ \sum^C_{j=1} \exp(\bm{f}_j)} \in \R^C$. If we write the targets $y_i$ as a one-hot $C$-shaped one-hot encoding vector, we can write the negative log likelihood as:
\begin{equation}
\begin{aligned}
	\ell(h(\bm{f}),\vy) &= - \log\left( \vy^\T \textrm{Softmax}(\bm{f})\right)\\
	&= - y_i \cdot \log \left( \left[ \textrm{Softmax}(\bm{f}) \right]_i \right)
\end{aligned}
\end{equation}
where $\left[\textrm{Softmax}(\bm{f})\right]_i = \frac{\exp(\bm{f}_i)}{\sum_{j=1}^{C} \exp(\bm{f}_j)} \in \R$ defines the $i$\textsuperscript{th} entry of the logits fed through the $\textrm{Softmax}(\cdot)$ function. Finally, we can write the first and second derivatives as follows

\begin{equation}
\begin{aligned}
	\nabla_f \ell(h(f), y_i)|_{f=f_i} &= - y_i + \left[ \textrm{Softmax}(\bm{f}) \right]_i \\
	\nabla_{f\!f} \ell(h(f), y_i)|_{f=f_i} &= \left[\textrm{Softmax}(\bm{f})\right]_i (1 - \left[\textrm{Softmax}(\bm{f})\right]_i).
\end{aligned}
\end{equation}

@@

%\begin{table}[!ht]
%\centering
%\caption{}
%\begin{tabular}{@{}cccc@{}}
%\toprule
%Loss        & $h(\cdot)$ & $\nabla_{f}$ & $\nabla^2_{f\!f}$  \\ \midrule
%Bernoulli   & $\sigma(f) = \frac{1}{1 + e^{-f}}$ & $\sigma(f) - y$   & $\sigma(f)(1-\sigma(f))$ \\
%MSE         & & $y - f$      & $1$ \\
%Categorical & $\mathcal{S}_i(\vf) = \frac{e^{\vf_i}}{\sum_{j=1}^C e^{\vf_j}}$ &  $\begin{bmatrix}
%		p_1\\ \vdots\\p_k
%	\end{bmatrix}$     &   $\begin{bmatrix}
%		p_1\\ \vdots\\p_k
%	\end{bmatrix}$ \\ \bottomrule
%\end{tabular}
%\end{table}
